---
title: "eds221-day10-activities"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
library(skimr)
```

Download the data

```{r}
snow <- read_csv(here::here("data", "ntl20_v6.csv"))
ice <- read_csv(here::here("data", "ntl33_v7.csv"), na = c("-999", ""))
```

Create a visualization that compares ice duration across different lakes in the Madison Lake Area

```{r}
ggplot(data = ice, aes(x = lakeid, y = lakeid )) +
  geom_jitter(width = 0.2, color = "gray50") +
  geom_boxplot(width = 0.3, fill = NA, size = 1, outlier.color = NA)
```

-   Find the mean ice cover duration in Madison Lake Area by year (i.e., you'll only be grouping by the year variable here). Understanding check: what are you finding the mean of here? Why are there multiple observations for each year?)

```{r}

mean_duration <- ice %>% 
  group_by(year4) %>% 
  summarize(mean_duration = mean(ice_duration, na.rm = TRUE))

```

-   Create an exploratory visualization of mean ice cover duration by year for the Madison Lake Area. Add a sentence or two below this exploratory graph describing the overall trend(s) you observe.

```{r}
ggplot(data = mean_duration, aes(x = year4, y = mean_duration)) +
  geom_line(color = "blue") + 
  ylim(min = 0, max = 200) +
  geom_abline()
```

Explanation of the graph: Slightly flutating to lower mean duration, the ice over time is slightly decreasing

-   Find the mean air temperature (using the daily average air temperature - adjusted) in Madison Lake Area **by year** using **only observations from winter months (December, January, February)**

```{r}
mean_air_temp <- snow |> 
  filter(month %in% c(1,2,12)) |> 
  group_by(year4) |> 
  summarize(mean_temp = mean(ave_air_temp_adjusted, na.rm = TRUE))
```

-   Create an exploratory visualization of mean winter temperatures for the Madison Lake Area. Add a sentence or two below this exploratory graph describing the overall trend(s) you observe.

```{r}
reg <- lm(mean_temp~year4, data = mean_air_temp)
ggplot(mean_air_temp, aes(x= year4, y=mean_temp)) +
  geom_point() +
  ylim(-15,5) +
  abline(reg, col = "red")

```

-   Join the mean winter air temperatures to the mean ice cover duration data you found above

```{r}

ice_weather_join <- mean_duration |> 
  full_join(mean_air_temp)


```

-   Create an exploratory scatter plot of mean winter temperatures versus mean ice duration. Add a sentence or two below this exploratory graph describing the overall trend(s) you observe.

```{r}
ggplot(data = ice_weather_join, aes(x = mean_temp, y = mean_duration)) +
  geom_jitter(aes(color = year4)) +
  scale_color_gradient(low = "blue", high = "red")
```

-   Explore thaw dates since 1970. Create a subset of the ice cover duration data since 1970 (ok to include 1970), then convert the `ice_off` column to a `Date`. Use `lubridate::yday()` to pull the numeric *day of the year* from that Date. Create an exploratory plot of the ice off day (numeric day-of-the-year) from 1970 - present. Add a sentence or two below this exploratory graph describing the overall trend(s) you observe.
